The Astropy Project

### 2017-11-13 main.tex
reorganize astropy.io

The \package{astropy.table} subpackage provides functionality for representing
and manipulating heterogeneous data.
\inlinecomment{APW:}{Need more introduction to the table subpackage!}
The package was described in detail in \cite{astropy}; we summarize key new
features or updates to \package{astropy.table} below.

\begin{description}
    \item[Support for grouped table operations] Some tables contain data that
        naturally form groups.
        For example, a table of photometry for galaxies, stars, and quasars.
        We may then want to split the table or a column into these groups and
        compute some statistic on the data, or filter the groups based on some
        criteria.
        Grouped operations are now supported by \texttt{Table} objects,
        mainly through the \texttt{group\_by()} method and subsequent operations
        on the grouped version of the table or column.
    \item[Support for table concatenation] Tables can be combined in several
        different ways.
        If two tables have the same rows, we may want to stack them
        ``vertically'' to create a new table with the same columns but all rows.
        If two tables are row-matched but have distinct columns, we may want to
        stack them ``horizontally'' to create a new table with the same rows
        but all columns.
        For other situations, more general table concatenations or joins are
        also possible when two tables share some common columns.
        All three of these operations are now possible with \texttt{Table}
        objects through the \texttt{vstack}, \texttt{hstack}, and \texttt{join}
        functions in \package{astropy.table}.
    \item[Mixin columns for time and coordinates] The default data containers
        for \texttt{Table} objects are \texttt{Column} objects.
        However, sometimes the data for a given column is better represented as
        a \texttt{Quantity}, or a sky coordinate (\texttt{SkyCoord}), or an
        astronomical \texttt{Time} object.
        The \texttt{Table} object now supports these and other objects to serve
        as array-like containers for column data through the ``mixin'' protocol.
    \item[Transform to and from \texttt{Pandas} \texttt{DataFrame} objects]
        \texttt{Table} objects can now be created from a \texttt{DataFrame}
        object (\texttt{from\_pandas()}), or exported to a \texttt{DataFrame}
        object (\texttt{to\_pandas()}).
        \texttt{DataFrame} objects do not support multi-dimensional array
        columns, so this conversion will fail if the source \texttt{Table}
        object contains any such columns.
\end{description}

@@

The \package{astropy.io} subpackage provides support for reading and writing
various formats: a wide range of ASCII data table formats, FITS, VOTable, etc.
It also provides a unified interface for reading and writing data with these
different formats, using the \package{astropy.table} subpackage.
For many common cases this simplifies the process of file I/O and reduce the
need to master the separate details of all the I/O packages within Astropy.
% This functionality is still in active development and the number of supported
% formats will be increasing.

@@

\subsubsection{Key new features}

\begin{description}

    \item[ASCII: fast reader/writer]

@@

    \item[ASCII: HTML tables]

@@

	\item [FITS: command-line scripts (\texttt{fitsheader}, \texttt{fitsinfo})]

@@

	\item [FITS: lazy loading ]

@@

	\item [Misc: YAML serialization ]

>>>>

The \package{astropy.table} subpackage provides functionality for representing
and manipulating heterogeneous data.
\inlinecomment{APW:}{Need more introduction to the table subpackage!}
The package was described in detail in \cite{astropy}.

We summarize key new features or updates to \package{astropy.table} below:
\begin{description}
    \item[Support for grouped table operations] Some tables contain data that
        naturally form groups.
        For example, a table of photometry for galaxies, stars, and quasars.
        We may then want to split the table or a column into these groups and
        compute some statistic on the data, or filter the groups based on some
        criteria.
        Grouped operations are now supported by \texttt{Table} objects,
        mainly through the \texttt{group\_by()} method and subsequent operations
        on the grouped version of the table or column.
    \item[Support for table concatenation] Tables can be combined in several
        different ways.
        If two tables have the same rows, we may want to stack them
        ``vertically'' to create a new table with the same columns but all rows.
        If two tables are row-matched but have distinct columns, we may want to
        stack them ``horizontally'' to create a new table with the same rows
        but all columns.
        For other situations, more general table concatenations or joins are
        also possible when two tables share some common columns.
        All three of these operations are now possible with \texttt{Table}
        objects through the \texttt{vstack}, \texttt{hstack}, and \texttt{join}
        functions in \package{astropy.table}.
    \item[Mixin columns for time and coordinates] The default data containers
        for \texttt{Table} objects are \texttt{Column} objects.
        However, sometimes the data for a given column is better represented as
        a \texttt{Quantity}, or a sky coordinate (\texttt{SkyCoord}), or an
        astronomical \texttt{Time} object.
        The \texttt{Table} object now supports these and other objects to serve
        as array-like containers for column data through the ``mixin'' protocol.
    \item[Transform to and from \texttt{Pandas} \texttt{DataFrame} objects]
        \texttt{Table} objects can now be created from a \texttt{DataFrame}
        object (\texttt{from\_pandas()}), or exported to a \texttt{DataFrame}
        object (\texttt{to\_pandas()}).
        \texttt{DataFrame} objects do not support multi-dimensional array
        columns, so this conversion will fail if the source \texttt{Table}
        object contains any such columns.
\end{description}

@@

The \package{astropy.io} subpackage provides support for reading and writing
data to a variety of ASCII and binary file formats, such as a wide range of
ASCII data table formats, FITS, and VOTable.
It also provides a unified interface for reading and writing data with these
different formats using the \package{astropy.table} subpackage.
For many common cases this simplifies the process of file input/output and
reduces the need to master the separate details of all the I/O packages within
\astropypkg.
% This functionality is still in active development and the number of supported
% formats will be increasing.

@@

We summarize key new features or updates to \package{astropy.io} below,
organized by sub-subpackage:
\begin{description}

    \item[\texttt{.ascii}: fast reader/writer]

@@

    \item[\texttt{.ascii}: HTML tables]

@@

	\item [\texttt{.fits}: command-line scripts (\texttt{fitsheader}, \texttt{fitsinfo})]

@@

	\item [\texttt{.fits}: lazy loading ]

@@

	\item [\texttt{.misc}: YAML serialization ]
